\sysfull

### 2017-11-04 paper/intro.tex
Introduction rewritten, refliecting better the novelty and contribution

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}. While both approaches represent progress, each has limitations. New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}. New programming models and compilers require programmers to change behaviour~\cite{chain,ratchet}, inhibiting adoption. \todo{Solve: no logical connection between data consistency models and memory models}{Przemek} New memory models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}. Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).

\textbf{Static Task-based Model and its Inefficiency:} Recently two {\em task-based} programming and execution models~\cite{alpaca,chain} were proposed in which programmers statically decompose a program into a collection of tasks. Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures. The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent. An key challenge is that the length of a software task's execution is \emph{limited by the fixed total amount of energy} that a device can buffer in hardware. A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict. To illustrate this, we refer to Figure~\ref{fig:1} showing the execution time of two applications (the same application X, but divided into X and Y tasks as in~\cite{chain}) running on RFID antenna-powered Computational RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna distances (near---stable energy supply/far---high energy intermittency). At short distance: execution will takes too long caused by runtime cost of marshalling excessively small tasks. At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer.

\textbf{Research Question and Contributions:} We need to know how to define and execute tasks efficiently and ensure non-termination, with an unknown energy buffer size and energy arrival process. In particular, we ask (i) how to use software support to \emph{dynamically adapt} the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while \emph{automatically maintaining memory consistency} during execution?

To address this problem, in this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}. To accomplish this, \sys introduces new capabilities:

\begin{itemize}
	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks as a single task. A programmer or compiler can specify small tasks that will execute on a device with a small energy buffer. Coalescing tasks allows running the same tasks on a device with a larger energy buffer, avoiding run time overheads associated with ending one task and beginning another.
	%
	\item {\bf Volatile memory virtualization:} \sys's memory virtualization uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution. When a task ends, \sys commits dirty pages to FRAM using DMA block copies, ensuring task atomicity and memory consistency.
\end{itemize}

As a helpful feature, in this work we also address the problem of automatic task compilation: \sys's compiler transforms arbitrary C code into a graph of tasks that reflects the original program's control- and data-flow constraints. The compiler identifies data shared by multiple of the newly created tasks and instruments reads and writes of those data so that \sys's memory virtualization mechanism keeps those data consistent. \sys's complier eliminates the labour-intensive process of writing tasks~\cite{chain,alpaca} (i.e., hours of programmer time).

We evaluated \sys end-to-end on a collection of existing benchmarks ~\cite{chain} and new programs, and compare directly to a state-of-the-art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexibly target a variety of platforms without rewriting code.

>>>>

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}. New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now). However, they have their own limitations that need to be addressed---the core one being inability to adapt to changing energy conditions.

\textbf{Static Task-based Model and its Inefficiency:} The latest two {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks. Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures. The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent. A key challenge is that the length of a software task's execution is \emph{limited by the fixed total amount of energy} that a device can buffer in hardware. A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict. To illustrate this, we refer to Figure~\ref{fig:1} showing the execution time of two applications (the same application X, but divided into X and Y tasks as in~\cite{chain}) running on RFID antenna-powered Computational RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna distances (near---stable energy supply/far---high energy intermittency). At short distance: execution will takes too long caused by runtime cost of marshalling excessively small tasks. At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer. This calls for at-runtime task division of any transiently-powered application.

\textbf{Data Inconsistency of Virtual Task:} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task. This however induces a problem of \emph{virtual task write-after-read (WAR) dependency violation}. Consider the following example of Task 1: {\tt{\{y=x+1;\}}}, Task 2: {\tt{\{z=y+1;\}}}, and Task 3 {\tt{\{x=z+1;\}}} being executed consecutively. All tasks have no WAR dependency and are consistent. Now, suppose three tasks have been virtualized into one, namely Task 4: {\tt{\{y=x+1; z=y+1; x=z+1;\}}} which has inherent WAR dependency (variable {\tt{x}}).

\textbf{Trade-off with Virtual Memory Manipulation:} WAR dependencies cannot be complied at runtime for merged (virtual) tasks. Therefore virtual task execution model needs to manipulate merged tasks in-memory to guarantee data consistency. Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}. Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}). How to connect this trade-off with task virtualization remains an open question.

\textbf{Research Question and Contributions:} Specifically, we ask (i) how to use software support to \emph{dynamically adapt} the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while \emph{automatically maintaining memory consistency} during execution?

To address this question, in this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}. To accomplish this, \sys introduces two new capabilities:

\begin{itemize}
	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks as a single task. A programmer or compiler can specify small tasks that will execute on a device with a small energy buffer. Coalescing tasks allows running the same tasks on a device with a larger energy buffer, avoiding run time overheads associated with ending one task and beginning another.
	%
	\item {\bf Volatile memory virtualization on transient power:} Colascing through task virtualization requires memory virtualization. \sys's uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution. When a task ends, \sys commits dirty pages to FRAM using DMA block copies, ensuring task atomicity and memory consistency. To the best of our knowledge this is the first time when memory virtualization has been demonstrated on a transient power budget.
\end{itemize}

We evaluated \sys end-to-end on a collection of existing benchmarks ~\cite{chain} and new programs, and compare directly to a state-of-the-art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexibly target a variety of platforms without rewriting code. \todo{Provide concrete numeric results supporting this claim}{Amjad/Przemek}

In this work we also address the problem of task construction burden imposed on the programmer. That is, we introduce \sys's compiler that transforms arbitrary C code into a graph of tasks that reflects the original program's control- and data-flow constraints. The compiler identifies data shared by multiple of the newly created tasks and instruments reads and writes of those data so that \sys's memory virtualization mechanism keeps those data consistent. While \sys's complier is still \emph{limited in scope}, e.g. is unable to handle instruction jumps, normalized loops and recursions, it already eliminates the labour-intensive process of writing tasks~\cite{chain,alpaca}, i.e., hours of programmer time.
